`r if(knitr:::is_latex_output()) '\\appendix'`

`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'` 

<!--
If you feel it necessary to include an appendix, it goes here.
-->

# I

<!-- This first appendix includes all of the R chunks of code that were hidden throughout the document (using the `include = FALSE` chunk tag) to help with readibility and/or setup. -->

<!-- **In the main Rmd file** -->

<!-- ```{r ref.label='include_packages', results='hide', echo = TRUE} -->
<!-- ``` -->

```{r yield-comparison-test, echo = FALSE}

# model_yield1 <- purrr::map(.x = paste(c("yield",
#                                        "effective_tiller", "thou_grain_wt",
#                                        "grain_per_pan", "panicle_len"), "7", sep = ""),
#                           .f = ~ lme4::lmer(get(.x) ~
#                                               (1 | rowgroup) +
#                                               (1 | colgroup) +
#                                               (1 | row:rowgroup) +
#                                               (1 | col:colgroup) +
#                                               (1 | interaction_rc) +
#                                               (1 | line_or_check),
#                                             data = all_data))
# 
# model_yield2 <- purrr::map(.x = paste(c("yield",
#                                        "effective_tiller", "thou_grain_wt",
#                                        "grain_per_pan", "panicle_len"), "7", sep = ""),
#                           .f = ~ lme4::lmer(get(.x) ~
#                                               (1 | rowgroup) +
#                                               (1 | colgroup) +
#                                               (1 | row:rowgroup) +
#                                               (1 | col:colgroup) +
#                                               (1 | interaction_rc),
#                                             data = all_data))

```


```{r leaf-health-comparison-test, echo=FALSE}

# model_leaf_health1 <- purrr::map(.x = c("dis_march15", "greenness_z65",
#                                        "greenness_z75", "greenness_z85",
#                                        "LAUG_65_75", "LAUG_75_85"),
#                                 .f = ~ lme4::lmer(get(.x) ~
#                                             (1 | rowgroup) +
#                                             (1 | colgroup) +
#                                             (1 | row:rowgroup) +
#                                             (1 | col:colgroup) +
#                                             (1 | interaction_rc) +
#                                             (1 | line_or_check),
#                                           data = all_data))
# 
# model_leaf_health2 <- purrr::map(.x = c("dis_march15", "greenness_z65",
#                                        "greenness_z75", "greenness_z85",
#                                        "LAUG_65_75", "LAUG_75_85"),
#                                 .f = ~ lme4::lmer(get(.x) ~
#                                             (1 | rowgroup) +
#                                             (1 | colgroup) +
#                                             (1 | row:rowgroup) +
#                                             (1 | col:colgroup) +
#                                             (1 | interaction_rc),
#                                           data = all_data))

```


```{r other-comparison-test, echo=FALSE}

# model_other_1 <- purrr::map(.x = c("plht", "lar",
#                                        "chl_ind1_2", "chl_ind2_2", "ctd",
#                                        "days_to_head", "days_to_anth"),
#                           .f = ~ lme4::lmer(get(.x) ~
#                             (1 | rowgroup) +
#                             (1 | colgroup) +
#                             (1 | row:rowgroup) +
#                             (1 | col:colgroup) +
#                             (1 | interaction_rc) +
#                             (1 | line_or_check),
#                           data = all_data))
# 
# model_other_2 <- purrr::map(.x = c("plht", "lar",
#                                        "chl_ind1_2", "chl_ind2_2", "ctd",
#                                        "days_to_head", "days_to_anth"),
#                           .f = ~ lme4::lmer(get(.x) ~
#                             (1 | rowgroup) +
#                             (1 | colgroup) +
#                             (1 | row:rowgroup) +
#                             (1 | col:colgroup) +
#                             (1 | interaction_rc),
#                           data = all_data))
```


**Tables referenced in Chapter \@ref(modeling-yield-and-yield-component-traits):**

```{r, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
# # save and load all previous chunk results
# save(list = c("model_yield1", "model_yield2", 
#               "model_leaf_health1", "model_leaf_health2", 
#               "model_other_1", "model_other_2"), 
#      file = "08-appendix.RData", 
#      envir = .GlobalEnv)
load("08-appendix.RData")

yield_models_compare <- pmap(list(model_yield, model_yield1, model_yield2),
                             .f = function(x, y, z) anova(x, y, z) %>%
                               broom::tidy() %>%
                               mutate(term = c("Blocking factors",
                                               "Blocking factors + Entry genotypes",
                                               "Blocking factors + Check and entry genotypes")))

pwalk(.l = list(yield_models_compare, 
               c("Yield", "Number of effective tillers", 
                 "Thousand kernel weight", "Number of grains per panicle", 
                 "Panicle length"), 
               c("lrt-yield", "lrt-netill", "lrt-tkw",
                 "lrt-gperpan", "lrt-panlen")),
     .f = function(x, y, z){knitr::kable(x, format = "latex", 
                                         booktabs = TRUE, align = "l", 
                                         caption = paste("\\label{tab:", z, "}", 
                                                         'ANOVA table with likelihood ratio test for ', y, sep = "")) %>% 
         kableExtra::kable_styling(latex_options = "HOLD_position") %>%
         kableExtra::column_spec(1, width = "3.5cm") %>% 
         print()
     })
```


**Tables referenced in Chapter \@ref(modeling-leaf-health-traits):**

```{r echo=FALSE, results='asis', warning=FALSE, message=FALSE}

leaf_health_models_compare <- pmap(list(model_leaf_health, model_leaf_health1, model_leaf_health2),
                                   .f = function(x, y, z) anova(x, y, z) %>% 
                                     broom::tidy() %>% 
                                     mutate(term = c("Blocking factors", 
                                                     "Blocking factors + Entry genotypes", 
                                                     "Blocking factors + Check and entry genotypes"))) 

pwalk(.l = list(leaf_health_models_compare, 
               c("Disease score at Zadok's stage 75", 
                 "Flag leaf greenness score at Zadok's stage 65", 
                 "Flag leaf greenness score at Zadok's stage 75", 
                 "Flag leaf greenness score at Zadok's stage 85", 
                 "LAUG between Zadok's stage 65 and 75", 
                 "LAUG between Zadok's stage 75 and 85"), 
               c("lrt-dis-score", "lrt-greenness1", "lrt-greenness2", "lrt-greenness3", "lrt-laug1", "lrt-laug2")),
     .f = function(x, y, z){knitr::kable(x, format = "latex", 
                                         booktabs = TRUE, align = "l", 
                                         caption = paste("\\label{tab:", z, "}", 
                                                         'ANOVA table with likelihood ratio test for ', y, sep = "")) %>% 
         kableExtra::kable_styling(latex_options = "HOLD_position") %>%
         kableExtra::column_spec(1, width = "3.5cm") %>% 
         print()
     })
```


**Tables referenced in Chapter \@ref(modeling-plant-morphological-architectural-and-phenological-traits):**

```{r echo = FALSE, results='asis', warning=FALSE, message=FALSE}

other_models_compare <- pmap(list(model_other, model_other_1, model_other_2),
     .f = function(x, y, z) anova(x, y, z) %>% 
       broom::tidy() %>% 
  mutate(term = c("Blocking factors", 
                  "Blocking factors + Entry genotypes", 
                  "Blocking factors + Check and entry genotypes"))) 

pwalk(.l = list(other_models_compare, 
                c("Plant height", 
                  "Flag leaf surface area", 
                  "SPAD at Zadok's stage 65", 
                  "SPAD at Zadok's stage 75", 
                  "Canopy temperature depression (CTD)", 
                  "Days to heading", 
                  "Days to anthesis"), 
                c("lrt-plht", "lrt-lar", 
                  "lrt-spadi", "lrt-spadii", 
                  "lrt-ctd", "lrt-dth", "lrt-dta")),
      .f = function(x, y, z){knitr::kable(x, format = "latex", 
                                          booktabs = TRUE, align = "l", 
                                          caption = paste("\\label{tab:", z, "}", 
                                                          'ANOVA table with likelihood ratio test for ', y, sep = "")) %>% 
          kableExtra::kable_styling(latex_options = "HOLD_position") %>%
          kableExtra::column_spec(1, width = "3.5cm") %>% 
          print()
      })
```


# II

\textbf{Snapshot of all statistical packages and processing environment used in analysis and interpretation of data:}

```{r echo=FALSE}
sessionInfo()
```

