
{\pagestyle{empty}
  \fontsize{12}{14}\selectfont
  \begin{titlepage}
  \newpage
  \let\footnotesize\small
  \let\footnoterule\relax
  \let \footnote \thanks

  \begin{center}
    \setcounter{page}{1}
    \null
    % \vfil
	% \topskip=5\baselineskip
	% \bigskip
    {\fontsize{12}{14}\selectfont \textbf{STUDY OF LEAF HEALTH AS POTENTIAL DETERMINANT OF YIELD IN EARLY EVALUATION OF DIVERSE WHEAT GENOTYPES IN RAMPUR, CHITWAN}}
	% \bigskip
    % \vskip 1.5em
    \vspace{8\baselineskip}
    % \bigskip
    \centerline{}
	{\fontsize{12}{14}\selectfont \lineskip .75em
    \begin{tabular}[t]{c}%
      DEEPENDRA DHAKAL
    \end{tabular}\par}
	\vspace{3\baselineskip}
    THESIS \\
    SUBMITTED TO THE \\
    AGRICULTURE AND FORESTRY UNIVERSITY \\
    DEPARTMENT OF GENETICS AND PLANT BREEDING \\
    RAMPUR, CHITWAN \\
    NEPAL \\
	\vspace{3\baselineskip}
	  IN PARTIAL FULFILMENT OF THE \\
	  REQUIREMENTS FOR THE \\
	  DEGREE OF \\
	\vspace{3\baselineskip}
	  MASTER OF SCIENCE IN AGRICULTURE \\
	  (PLANT BREEDING AND GENETICS) \\
    \centerline{}
    {\fontsize{12}{14}\selectfont 2018 \par}
  \end{center}\par
  \end{titlepage}
}

\clearpage
  {\fontsize{12}{14}
  \centerline{\bfseries CERTIFICATE}
  \thispagestyle{empty}
  \null\vspace*{2cm}

  \begingroup
  This is to certify that the thesis/dissertation entitled \textbf{``STUDY OF LEAF HEALTH AS POTENTIAL DETERMINANT OF YIELD IN EARLY EVALUATION OF DIVERSE WHEAT GENOTYPES IN RAMPUR, CHITWAN''} submitted in partial fulfillment of the requirements for the degree of \textbf{MASTER OF SCIENCE IN AGRICULTURE} (\textbf{PLANT BREEDING AND GENETICS}) of the Agriculture and Forestry University, Nepal is a record of original research carried out by Mr. \textbf{Deependra Dhakal} Id. No. \textbf{PLB-03M-2015}, under my supervision, and no part of the thesis/dissertation has been submitted for any other degree or diploma.
  \endgroup
  
  \vskip 1in		%space to sign
  
  \noindent {\rule{3.5cm}{.4pt}\\
Dr. Madhav Prasad Pandey\\
(Chairman of the Advisory Committee)\\
Department of Genetics and Plant Breeding\\
Date:  \\[0.75cm]}
  
  \clearpage
}

\clearpage
  {\fontsize{12}{14}
  \centerline{\bfseries CERTIFICATE}
  \thispagestyle{empty}
  \null\vspace*{2cm}

  \begingroup
  We, the undersigned, members of the Advisory Committee of Mr. \textbf{Deependra Dhakal} Id. No. \textbf{PLB-03M-2015}, a candidate for the degree of \textbf{MASTER OF SCIENCE IN AGRICULTURE (PLANT BREEDING AND GENETICS)} agree that the thesis entitled \textbf{``STUDY OF LEAF HEALTH AS POTENTIAL DETERMINANT OF YIELD IN EARLY EVALUATION OF DIVERSE WHEAT GENOTYPES IN RAMPUR, CHITWAN''} may be submitted in partial fulfillment of the requirements for the degree.
  \endgroup
  
  \vskip 1cm		%space to sign
  
  \begingroup
  
	{
  \setbox0=\hbox{\makebox[3.5cm]{}}
  \thispagestyle{empty}
  \vskip 1cm		%space to sign
  {\makebox[\wd0][l]{\hrulefill}
	\makebox[1.5in]{} \makebox[\wd0][l]{\hrulefill}}\\
  {\makebox[\wd0][l]{Dr. Madhav Prasad Pandey}
	\makebox[1.5in]{} \makebox[\wd0][l]{Mr. Nutan Raj Gautam}}\\
	{\makebox[\wd0][l]{(Member of the Advisory Committee)}
	\makebox[1.5in]{} \makebox[\wd0][l]{(Scientist)}}\\
	{\makebox[\wd0][l]{Department of Genetics and Plant Breeding}
	\makebox[1.5in]{} \makebox[\wd0][l]{National Wheat Research Program}}\\
	{\makebox[\wd0][l]{Date:  }
	\makebox[1.5in]{} \makebox[\wd0][l]{Date:  }}
  % \par\vfil\null
  \vspace{1.0cm}
  }
  
  	\noindent {\rule{3.5cm}{.4pt}\\
Mr. Bishnu Raj Ojha\\
(Member of the Advisory Committee)\\
Department of Genetics and Plant Breeding\\
Date:  \\[0.75cm]}
	
	% this works too but for single advisor only
	% \raggedright
  % \makebox[\textwidth][l]{Dr. Madhav Prasad Pandey} \\
  % \makebox[\textwidth][l]{(Chairman of the Advisory Committee)} \\
  % \makebox[\textwidth][l]{Department of Genetics and Plant Breeding} \\
  % \makebox[\textwidth][l]{Date: \today }
  % \vskip 1cm		%space to sign
  \endgroup
  
  \clearpage
}
